Reviewer Comments (Scientific Reports)

### 2020-02-03 notebook/sci_rep_reviews.Rmd
Address Reviewer 2 comments

@@

@@

As Reviewer 2 correctly points out, my focus is on finite values of S and \sigma, in which complexity is gradually increased until it is dangerously close to the stability boundary. Given this focus, analytically demonstrating my central result (that variation in $X$ when $M = XA$ increases stability) would require some sort of comparison between the probability distributions of eigenvalues for $M$ and $A$. For circular matrices in which elements are i.i.d ($A$), it has been proved that eigenvalues are distributed uniformly within a known boundary of support of the spectral density (Tao and Vu 2010). While recent papers have found the support for more general types of random matrices that include the form $M = AX$ (e.g, Gibbs et al. 2018; Neri and Metz 2012; Ahmadian et al. 2015), to the best of my understanding, finding the actual density of eigenvalues within the support has not been attempted. This likely matters because for finite S, even though the real boundary of the support might be positive for both $M$ and $A$ (implying instability as $S \to \infty$; Grilli et al. 2017), the right most eigenvalue for a system still might be less than zero.

My approach follows others with similar goals (Sinha 2005).

Others have found the support

Reviewer 2 is correct that analytical results here are uncharted territory, despite

In recent papers, analytical solutions have been provided to find the support of the spectral densities of matrices of the form $M = XA$,

This is unique, as most studies make the simplifying assumption of \sigma = \sqrt{S} such that the circular matrix has a radius of unity.

Ideally, an analytical proof could demonstrate the uncertainty in the position of the leading eigenvalue at finite S, which would likely require knowledge of the distribution of M = \gamma A for arbitrary values of \sigma and S.

For matrices of the form $M = XA$, Gibbs et al. (2018) do find the boundaries of support for spectral densities.

My focus is uniquely on a finite S. This is important, as Ahmadian et al. (2015) showed that any realisation of the eigenvalue distribution for a matrix of the form Z = M + LJR approaches the expected distribution as S \to infty.

Elements of M = XA are no longer independent.

@@

@@

>>>>

Author response to Reviewer 2 comment 1
================================================================================

@@

Author response to Reviewer 2 comment 2
================================================================================

@@

As Reviewer 2 correctly points out, my focus is on finite values of S and \sigma, in which complexity is gradually increased until it is dangerously close to the stability boundary. Given this focus, analytically demonstrating my central result (that variation in $X$ when $M = XA$ increases the stability of $M$), would require some sort of comparison between the probability distributions of eigenvalues for $M$ and $A$. For circular matrices in which elements are i.i.d ($A$), it has been proved that eigenvalues are distributed uniformly within a known boundary of support of the spectral density (Tao and Vu 2010). While recent papers have found the support for more general types of random matrices that include the form $M = AX$ (e.g, Gibbs et al. 2018; Neri and Metz 2012; Ahmadian et al. 2015), to the best of my understanding, finding the actual density of eigenvalues within the support has not been attempted. This likely matters because for finite S, even though the real boundary of the support might be positive for both $M$ and $A$ (implying instability as $S \to \infty$; Grilli et al. 2017), the right most eigenvalue for a system still might be less than zero.

Reviewer 2 is correct that this is uncharted territory, and after carefully considering the possibility of analytical demonstration of my manuscript's results, I am convinced that such a demonstration is beyond the scope of this manuscript. My approach follows others with similar objectives. For example, Sinha (2005) also used a numerical approach to understand the stability of small-world networks (in contrast to the classic random networks of May).

Reviewer 2 comment 3
================================================================================

@@

Author response to Reviewer 2 comment 3
================================================================================

These were very helpful comments, and I am especially grateful to Reviewer 2 for doing a bit of preliminary exploration themselves to suggest improvements. I have now run simulations for different values of $\rho$, as suggested. Reviewer 2 is correct that the nonzero $\rho$ values do not change my results, but an analysis of stability across $\rho$ values also revealed an interesting pattern. While the relationship between $\rho$ and the leading eigenvalue is linear for circular matrices ($A$), it is convex when $\gamma$ is included such that $M = \gamma A$. I have now included these results into the main text of the manuscript, as I believe that they are important.

As suggested, I also considered matrices with even more structure. Specifically, I have simulated three broad classes of structured networks that are widespread in physical, social, and biological systems. These include food webs with cascade structures (e.g., Williams and Martinez 2000), small-world networks (Watts and Strogatz 1998), and scale-free networks (Albert and Barab\'{a}si 2002).

Reviewer 2 comment 4
================================================================================

@@

Author response to Reviewer 2 comment 4
================================================================================

While I agree that the effect of $var(\gamma)$ is restricted to a small proportion systems at the margins of stability, I would argue that this effect is actually of quite general relevance because of the breadth of complex systems to which my results apply. Addressing Reviewer 2's previous comment helped emphasise this; the effect of variation in $\gamma$ on stability, while small in overall magnitude, is predicted to be general across any physical, social, or biological network. Since nearly all such networks will have components that vary in the rate at which they respond to perturbation, my results contribute to a general understanding of system stability.

I have revised the writing to more carefully frame the manuscript's key results (line XXX). This includes a change of the title, which now makes it more clear that the effects on stability are small while retaining an emphasis on generality across systems. I have also removed the word "markedly" from the abstract (line XXX).
